Alone Together: Compositional Reasoning and Inference for Weak Isolation

### 2017-10-19 inference.tex
Full encoding algorithm

The algorithm ($\mssemof{\cdot}$) shown in Fig.~\ref{fig:logic}
encodes an $\SL$ expression in a decidable fragment of FOL. Given a
set expression $s$ of $\SL$, $\mssemof{s}$ is a pair $(\phi,f)$, where
$\phi$ encodes $s$ in FOL, and $f$ is a Boolean function that names
the set defined by $s$.  Naming provides an easy handle to $s$,
simplifying $\mssemof{\cdot}$.  Set variables ($x$, $\stl$, $\stg$,
$\dots$) are encoded as unconstrained Boolean functions ($g_x$,
$g_{\stl}$, $g_{\stg}$, $\ldots$) that are named after the variable.
Encoding of a set literal $\{a\,|\,\varphi\}$ introduces a new Boolean
function $g$ that is \emph{true} for only those $a$ that satisfy $\varphi$.
Thus $g$ faithfully encodes the literal. The encoding of the
$\existsl{}$ expression first eliminates the second-order
quantification over sets.  Existentially bound set variable ($a$) is
skolemized, i.e., instantiated with a new unbound variable $a'$,
before encoding the bound expression $s$ and the first-order
constraint $\phi$. The algorithm for encoding $\phi$, which is already
a first-order formula, is straightforward. It basically involves
replacing set variables with Boolean functions named after the
variables. The first-order encoding for a bind expression describes
the semantics of set monad's bind operator in FOL. Let $s_1$ be a set, and
let $f$ be a function that maps each variable in $s_1$ to a new
set. Then, $s_2 = s_1 \bind f$ if and only if for all $y\in s_2$,
there exists an $x \in s_1$ such that $y = f(x)$, and forall $x\in
s_1$, $f(x)\in s_2$. The encoding adds new constraints to this
effect. Conditional set expressions and set union are encoded
straightforwardly.

>>>>

The algorithm ($\mssemof{\nubar}{\cdot}$) shown in Fig.~\ref{fig:logic}
encodes an $\SL$ expression in a decidable fragment of FOL. Given a
set expression $s$ of $\SL$, $\mssemof{\nubar}{s}$ is a pair $(\phi,f)$, where
$\phi$ encodes $s$ in FOL, and $f$ is a Boolean function that names
the set defined by $s$.  Naming provides an easy handle to $s$,
simplifying $\mssemof{\nubar}{\cdot}$.  Set variables ($x$, $\stl$, $\stg$,
$\dots$) are encoded as unconstrained Boolean functions ($g_x$,
$g_{\stl}$, $g_{\stg}$, $\ldots$) that are named after the variable.
Encoding of a set literal $\{a\,|\,\varphi\}$ introduces a new Boolean
function $g$ that is \emph{true} for only those $a$ that satisfy $\varphi$.
Thus $g$ faithfully encodes the literal. The encoding of the
$\existsl{}$ expression first eliminates the second-order
quantification over sets.  Existentially bound set variable ($a$) is
skolemized, i.e., instantiated with a new unbound variable $a'$,
before encoding the bound expression $s$ and the first-order
constraint $\phi$. The algorithm for encoding $\phi$, which is already
a first-order formula, is straightforward. It basically involves
replacing set variables with Boolean functions named after the
variables. The first-order encoding for a bind expression describes
the semantics of set monad's bind operator in FOL. Let $s_1$ be a set, and
let $f$ be a function that maps each variable in $s_1$ to a new
set. Then, $s_2 = s_1 \bind f$ if and only if for all $y\in s_2$,
there exists an $x \in s_1$ such that $y = f(x)$, and forall $x\in
s_1$, $f(x)\in s_2$. The encoding adds new constraints to this
effect. Conditional set expressions and set union are encoded
straightforwardly.
